Is essentiality of genes conserved in Enterobacteriaceae?

### 2017-02-19 documents/manuscript.tex
Ran prokka on Enterobacter and reran everything, corrected parsimony code

Although there are some studies on differentiation of essentiality in different organisms, these studies usually include a few genomes. We are going to extend these studies to a larger scale in Enterobacteriaceae family. Our aim is to study the essentiality of genes in in 16 different organisms. These include \textit{Enterobacter cloacae} NCTC 9394, \textit{Klebsiella pneumoniae} Ecl8, \textit{Klebsiella pneumoniae} RH201207, \textit{Citrobacter rodentium} ICC168, \textit{Salmonella} Typhimurium SL1344, \textit{Salmonella} Typhimurium SL3261, \textit{Salmonella} Typhimurium D23580, \textit{Salmonella} Typhimurium A130, \textit{Salmonella} Enteritidis P125109, \textit{Salmonella} Typhi Ty2, \textit{Escherichia coli} ST131 EC958, \textit{Escherichia coli} UPEC ST131, \textit{Escherichia coli} ETEC CS17, \textit{Escherichia coli} ETEC H10407, \textit{Escherichia coli} BW25113, and \textit{Escherichia coli} K-12 MG1655. The phylogenetic tree containing all these species is depicted in Fig.\@ \ref{fig:species-tree}.

@@

\begin{figure}
\includegraphics[scale=0.2]{phylosift-aa-raxmlbootstrap.pdf}
\caption{The species tree containing the 16 genomes used in this study. We have generated the tree by running RAxML \cite{stamatakis_raxml_2014} on Phylosift \cite{darling_phylosift:_2014} amino acid markers.}
\label{fig:species-tree}
\end{figure}

@@

\begin{figure}
\captionsetup[subfigure]{justification=centering}
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[page=1, scale=0.4]{essential-call-comparison-ETEC.pdf}
  \caption{}
  \label{fig:benchmark}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[page=10, scale=0.4]{essential-call-comparison-ETEC.pdf}
  \caption{}
  \label{fig:pca}
\end{subfigure}
\caption{(a) The accuracy of 7 different prediction methods for quantifying the essentiality of genes. The higher the area under the curve, the more accurate the method is. (b) The distribution for our proposed method (NPEQ)}
\label{fig:essentiality-call}
\end{figure}

@@

\begin{figure}
\captionsetup[subfigure]{justification=centering}
\begin{subfigure}{.49\textwidth}
  \centering
  \includegraphics[page=26, scale=0.25]{biases.pdf}
  \caption{}
  \label{fig:distance-bias}
\end{subfigure}
\begin{subfigure}{.49\textwidth}
  \centering
  \includegraphics[page=25, scale=0.25]{biases.pdf}
  \caption{}
  \label{fig:GC-bias}
\end{subfigure}
\begin{subfigure}{\textwidth}
  \centering
  \includegraphics[scale=0.6]{100logo-prob-edited.pdf}
  \caption{}
  \label{fig:logos}
\end{subfigure}
\begin{subfigure}{.49\textwidth}
  \centering
  \includegraphics[scale=0.25, page=2]{insertion-position-bias.pdf}
  \caption{}
  \label{fig:insertion-position-bias-ess}
\end{subfigure}
\begin{subfigure}{.49\textwidth}
  \centering
  \includegraphics[scale=0.25, page=4]{insertion-position-bias.pdf}
  \caption{}
  \label{fig:insertion-position-bias-ben}
\end{subfigure}
\caption{(a)The distance of the genes from DnaA gene normalised by the lengths of the genomes versus NPEQ. The distance from DnaA gene has been calculated in both directions and then the minimum value has been used for distance. The red curves show the loess curve when the smoothness parameter is $0.2$ and the green line shows no bias. (b) G-C content of genes against their NPEQ values. The red curves show the loess curve when the smoothness parameter is $0.2$ and the green line shows no bias. (c) Sequence logo plots generated using sequences from 10 nucleotides flanking the 100 top most frequent insertion sites from each genome. The height of each character shows the relative frequency of that character. (d) and (e) The plots show the average insertion index in percentiles of all essential genes (d) and beneficial losses (e). The genes are divided into 3 segments: 5\% of the genes on the $5^\prime$ end, 20\% of the genes on the $3^\prime$ end, and the rest in the middle. These are shown by khaki, slate gray, and violet red respectively.}
\label{fig:biases}
\end{figure}

@@

\subsection{phylogenetic analysis of gene essentiality identifies ...}
%\section{Discussion}
\section{Materials and Methods}
%\subsection{Transposon insertion}
%\section{Results and discussion}
%Throughout time, species can gain or lose genes. We investigated if these gene gain and losses are related to the essentiality of the genes. In this section, we have first described the biases that can affect our study, and then evaluated the essentiality of genes and their conservation and the relationship between these two.
%
%
%%\subsection{Length bias}
%%Short genes are less probable to be hit by transposons and so the genes that are very short might have no insertions. As a result, normalising the number of insertions by the length of the genes does not affect the essentiality level of these genes. 
%%In Fig.\@ \ref{fig:length-bias} we have studied if the length of genes can influence their essentiality and we have found no correlation between length and insertion index in short genes. It seems that the insertion density in our experiment is high enough to make it possible for every gene to be hit by transposons.
%%\begin{figure*}
%%\centering
%%\includegraphics[page=42, scale=0.3]{biases.pdf}
%%\caption{The plot shows the lengths of the genes against their insertion indices. The red line shows the loess curve when the smoothness parameter is $0.2$.}
%%\label{fig:length-bias}
%%\end{figure*}
%
%
%\subsubsection{Preferred insertion motif bias}

>>>>

Although there are some studies on differentiation of essentiality in different organisms, these studies usually include a few genomes. We are going to extend these studies to a larger scale in Enterobacteriaceae family. Our aim is to study the essentiality of genes in in 16 different organisms. These include \textit{Enterobacter cloacae} NCTC 9394, \textit{Klebsiella pneumoniae} Ecl8, \textit{Klebsiella pneumoniae} RH201207, \textit{Citrobacter rodentium} ICC168, \textit{Salmonella} Typhimurium SL1344, \textit{Salmonella} Typhimurium SL3261, \textit{Salmonella} Typhimurium D23580, \textit{Salmonella} Typhimurium A130, \textit{Salmonella} Enteritidis P125109, \textit{Salmonella} Typhi Ty2, \textit{Escherichia coli} ST131 EC958, \textit{Escherichia coli} UPEC ST131, \textit{Escherichia coli} ETEC CS17, \textit{Escherichia coli} ETEC H10407, \textit{Escherichia coli} BW25113, and \textit{Escherichia coli} K-12 MG1655.

@@

%\begin{figure}
%\includegraphics[scale=0.2]{phylosift-aa-raxmlbootstrap.pdf}
%\caption{The species tree containing the 16 genomes used in this study. We have generated the tree by running RAxML \cite{stamatakis_raxml_2014} on Phylosift \cite{darling_phylosift:_2014} amino acid markers.}
%\label{fig:species-tree}
%\end{figure}

@@

\begin{figure}
\captionsetup[subfigure]{justification=centering}
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[page=1, scale=0.4]{essential-call-comparison-BW25113.pdf}
  \caption{}
  \label{fig:benchmark}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[page=10, scale=0.4]{essential-call-comparison-BW25113.pdf}
  \caption{}
  \label{fig:pca}
\end{subfigure}
\caption{(a) The accuracy of 7 different prediction methods for quantifying the essentiality of genes. The higher the area under the curve, the more accurate the method is. (b) The distribution for our proposed method (NPEQ)}
\label{fig:essentiality-call}
\end{figure}

@@

\begin{figure}
\captionsetup[subfigure]{justification=centering}
\begin{subfigure}{.49\textwidth}
  \centering
  \includegraphics[page=6, scale=0.25]{biases.pdf}
  \caption{}
  \label{fig:distance-bias}
\end{subfigure}
\begin{subfigure}{.49\textwidth}
  \centering
  \includegraphics[page=5, scale=0.25]{biases.pdf}
  \caption{}
  \label{fig:GC-bias}
\end{subfigure}
\begin{subfigure}{\textwidth}
  \centering
  \includegraphics[scale=0.6]{100logo-prob-edited.pdf}
  \caption{}
  \label{fig:logos}
\end{subfigure}
\begin{subfigure}{.49\textwidth}
  \centering
  \includegraphics[scale=0.25, page=2]{insertion-position-bias.pdf}
  \caption{}
  \label{fig:insertion-position-bias-ess}
\end{subfigure}
\begin{subfigure}{.49\textwidth}
  \centering
  \includegraphics[scale=0.25, page=4]{insertion-position-bias.pdf}
  \caption{}
  \label{fig:insertion-position-bias-ben}
\end{subfigure}
\caption{(a)The distance of the genes from DnaA gene normalised by the lengths of the genomes versus NPEQ. The distance from DnaA gene has been calculated in both directions and then the minimum value has been used for distance. The red curves show the loess curve when the smoothness parameter is $0.2$ and the green line shows no bias. (b) G-C content of genes against their NPEQ values. The red curves show the loess curve when the smoothness parameter is $0.2$ and the green line shows no bias. (c) Sequence logo plots generated using sequences from 10 nucleotides flanking the 100 top most frequent insertion sites from each genome. The height of each character shows the relative frequency of that character. (d) and (e) The plots show the average insertion index in percentiles of all essential genes (d) and beneficial losses (e). The genes are divided into 3 segments: 5\% of the genes on the $5^\prime$ end, 20\% of the genes on the $3^\prime$ end, and the rest in the middle. These are shown by khaki, slate gray, and violet red respectively.}
\label{fig:biases}
\end{figure}

@@

\subsection{phylogenetic analysis of gene essentiality identifies that essential genes are more likely to be conserved}

\begin{figure}
\captionsetup[subfigure]{justification=centering}
\begin{subfigure}{.6\textwidth}
  \centering
  \includegraphics[scale=0.12]{phylosift-aa-raxmlbootstrap-annotated.pdf}
  \caption{}
  \label{fig:tree}
\end{subfigure}
\begin{subfigure}{.4\textwidth}
  \centering
  \includegraphics[scale=0.4]{fitch.pdf}
  \caption{}
  \label{fig:fitch}
\end{subfigure}

%\begin{subfigure}{.5\textwidth}
%  \centering
%  \includegraphics[page=2, scale=0.45]{upsetr.pdf}
%  \caption{}
%  \label{fig:upsetr1}
%\end{subfigure}
%\begin{subfigure}{1\textwidth}
%  \centering
%  \includegraphics[page=3, scale=0.65]{upsetr.pdf}
%  \caption{}
%  \label{fig:upsetr2}
%\end{subfigure}
\caption{(a) The species tree for all genomes in this study. Numbers in red show the number of core essential genes at each level and numbers in blue show the number of core genes at each level. (b) The ratio between core essential genes and core genes at each level in the species tree. The line connects the medians in each level.}
\label{fig:essentiality-phylogeny}
\end{figure}
%\section{Discussion}
\section{Materials and Methods}
%\subsection{Transposon insertion}
%\section{Results and discussion}
%Throughout time, species can gain or lose genes. We investigated if these gene gain and losses are related to the essentiality of the genes. In this section, we have first described the biases that can affect our study, and then evaluated the essentiality of genes and their conservation and the relationship between these two.
%
%
%%\subsection{Length bias}
%%Short genes are less probable to be hit by transposons and so the genes that are very short might have no insertions. As a result, normalising the number of insertions by the length of the genes does not affect the essentiality level of these genes. 
%%In Fig.\@ \ref{fig:length-bias} we have studied if the length of genes can influence their essentiality and we have found no correlation between length and insertion index in short genes. It seems that the insertion density in our experiment is high enough to make it possible for every gene to be hit by transposons.
%%\begin{figure*}
%%\centering
%%\includegraphics[page=42, scale=0.3]{biases.pdf}
%%\caption{The plot shows the lengths of the genes against their insertion indices. The red line shows the loess curve when the smoothness parameter is $0.2$.}
%%\label{fig:length-bias}
%%\end{figure*}
%
%
%\subsubsection{Preferred insertion motif bias}
